### 2017-12-13 paperDraft3.tex
update

\documentclass[12pt]{article}
\usepackage{amsmath, amssymb, fullpage}
\usepackage{array}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{caption}
\usepackage{subcaption}

\begin{document}

\title{Bayesian Hierarchical Models for Common Components across Multiple System
Configurations}

\begin{abstract}
    We use a Bayesian hierarchical model to assess the reliability of a family
    of vehicles. The proposed model effectively combines information across
    three phases of testing and across common vehicle components. The analysis
    yields estimates of failure rates across failure modes and vehicles as well
    as an overall estimate of the failure rate for the family of vehicles. We
    are also able to obtain estimates of the Fix Effectiveness Factor (FEF), a
    measure of how well repairs between test phases improve failure rates. In
    addition to using all data to improve on current assessments of reliability
    and reliability growth, we illustrate how to leverage the information
    learned from the three developmental test phases to determine the length of
    testing needed to demonstrate a given reliability threshold in the next
    testing phase.
\end{abstract}

\section{Introduction}
Reliability is a high priority in the testing of military systems. A common
interest is {\em reliability growth}, which tracks the change, and ideally
improvement, in the reliability of a system as it moves through testing phases
and undergoes periodic corrective actions. Ideally, plans for future test events
should be based on what is observed in previous testing.

In this paper, we look at combining information across multiple test phases with
the objective of planning a future test. When combining information, there is no
omnibus solution. Rather, models need to be carefully considered and evaluated
to ensure that they accurately reflect the data and the underlying physical
processes. Our analysis is motivated by reliability data from the Joint Light
Tactical Vehicle (JLTV), which is a family of vehicles designed to replace
one-third of the legacy high-mobility multipurpose wheeled vehicle (Humvee)
fleet. There are four unique types of vehicles, and within types there are many
possible variants. The primary mission of the family of vehicles is to provide
ground mobility that is deployable worldwide and capable of operating across the
range of military roles including combat, sustainment, police action,
peacekeeping, and security patrol, in all weather and terrain conditions.

The JLTV has been through a series of test events as it has been developed.
Engineering and Manufacturing Development (EMD) included three phases of
testing, where fixes occurred only during a set Corrective Action Period (CAP)
between each phase. For every vehicle, each failure encountered during testing
was recorded and attributed to a specific failure mode.  Since fixes were
delayed to CAPs, analysis generally shows a distinct jump in the system
reliability between test phases.

We propose a model that can effectively combine information across the three
phases of testing and across common vehicle components. We use a Bayesian
hierarchical model to assess the reliability of the family of vehicles. The
model accounts for the commonality across vehicle types and allows for
uncertainty quantification. Inferential objectives from the proposed model
include an estimate of the mean miles between failure (MMBF) for each vehicle,
failure mode, and phase of test, and an assessment of the effectiveness of the
corrections completed in the CAPs. Additionally, because future tests are being
planned, we propose methodology to leverage the information learned from the
three phases to determine the length of testing needed to demonstrate a given
reliability threshold in subsequent testing.

\section{Data}
Eight vehicle prototypes were used for reliability, availability,
and maintainability (RAM) testing. There were four two-seat utility vehicles,
two four-seat general purpose vehicles, a two-seat heavy guns carrier, and a
two-seat close combat weapons carrier. The close combat weapons carrier, one
general purpose, and two utility vehicles were tested in Aberdeen, Maryland,
where it is much colder in the winter months but the terrain is not as harsh,
and the heavy weapons carrier, one general purpose, and two utility vehicles
were tested in Yuma, Arizona, where it is warmer but with more challenging
terrain, to attain testing in different environmental and terrain conditions.
The vehicles were driven roughly 3000 to about 5000 miles in Phase 1, between
4500 and 7500 miles in Phase 2, and between 5000 and 6700 miles in Phase
3.\footnote{The exact mileage has been altered to protect proprietary
information.}

There are many similarities between vehicle types; in fact, the most dissimilar
vehicles have over 80\% common parts. Consequently, some failure modes, such as
brakes and radios, will be common to all eight vehicles. Other failure modes
will be related, but not identical, as with hydraulics, frame, and body.

There are 1412 observed failures attributed to 26 failure modes recorded across
all eight vehicles and all three test phases. In any given phase of test, every
vehicle is not guaranteed to have a failure of all 26 failure modes, and each
test phase does not necessarily end with a failure. Therefore, we have 624 right
censored observations, accounting for miles driven without observed failures.

\section{Methodology}

\subsection{Modeling Reliability}
A standard reliability analysis employed by the Department of Defense (DoD) test
community considers each test phase independently and uses the exponential
distribution to model the miles between failure for the system ~\cite{ref1}. The
traditional analysis is overly simplistic, relies on correct modeling
assumptions and ignores valuable information learned about the individual
vehicles and their failure modes.

In this paper, we propose an alternative approach using a Bayesian hierarchical
model.  We will begin by introducing a hierarchical structure that models the
relationships in the data across all test phases and incorporates known
similarities between vehicle failure modes.  Next, we will consider different
distributional assumptions and their implications.  We will also
discuss model diagnostics to consider when choosing a final model. Then we will
use the JLTV data to illustrate this process and discuss system reliability
results. This will lead us into the next section on using these results to
improve the planning of future test.

In this reliability setting we are analysing time to failure data measured in
miles for a multiple component system with the primary interest in the system
failure rate and future failure counts.  The most common models used in this
failure rate analysis are lifetime distributions and the most common for failure
count analysis is the \emph{Poisson process} counting model.  We will utilize
both and the relationships between them.

The methodology section is organized as follows... section 3.1...

\subsubsection{Model Structure}
Introducing notation, for each phase of testing we have failure mileage data for
a given vehicle $i$ and failure mode $j$ denoted as $t_{ij0}, t_{ij1},...,
t_{ijn_{ij}}, t_{ijc}$, where $t_{ij0} = 0$ and $t_{ijc}$ is the end of that
testing phase. In each phase of testing, we assume the vehicle miles between
failures, $y_{ijk} = t_{ijk} - t_{ijk-1}$, follow a lifetime distribution with
an unknown set of parameters that include a rate parameter.  For the first phase
the rate parameter for vehicle $i$ and failure mode $j$ will be denoted as
$\lambda_{ij}$.  Then for the next stage we will use the same $\lambda_{ij}$ but
add a multiplicative correction parameter $\rho_{1,j}$ that represents the
change in this rate due to the CAP.  This proceedure is then repeated for each
subsequent stage.  The correction parameter will be discussed more in the next
section but the general model is as follows,
\begin{align*}
\text{Phase 1: }&y_{ijk}\mid\lambda_{ij}\sim f(\cdot), \quad i = 1,2,...,v \quad
j=1,2,...,s \quad k=1,2,...,n_{ij}\\
\text{Phase 2: }&y_{ijk}\mid\lambda_{ij}\rho_{j}^{P2}\sim f(\cdot) \\
\text{Phase 3: }&y_{ijk}\mid\lambda_{ij}\rho_{j}^{P2}\rho_{j}^{P3}\sim f(\cdot)
\end{align*}
where $v$ represents the number of vehicles, $s$ is the number of failure modes,
and $n_{ij}$ are the number of failures of vehicle $i$ failure mode $j$. The
number of failure modes is assumed fixed and known \textit{a priori}.

For this model we will assume a prior distribution on failure rate parameter
$\lambda_{ij}$.  This distribution depends on whether failure mode $j$ is
considered to be common across vehicles or related but not identical. For the
related failure modes, we place a prior distribution on the collection of
$\lambda_{ij}$; in other words, we assume each vehicle has a distinct failure
rate in failure mode $j$ but they arise from a common distribution.  If failure
mode $j$ is considered common across vehicles, the collection of failure rates
is collapsed to a single parameter, $\lambda_{ij} = \lambda_j$. As with the
related failure modes, a prior distribution is placed on the single failure
rate. The prior distributions are independent across failure mode, and can
potentially have different hyperparameters.

% not sure we need this paragraph %
The Phase 1 analysis yields an estimate of the failure rate $\lambda_{ij}$ for
each of the vehicles for failure modes that are related.  We are assuming the
vehicles are conditionally independent, if we also assume the miles at failure
come from an exponential distribution the overall failure rate across all
failure modes can be easily calculated by $\sum_{j}\lambda_{j}$.  The
exponential distribution is commonly used in reliability analysis because of its
convenience, but as we will discuss in later sections it can be unrealistic in
many real-world situations.

\subsubsection{Fix Effectiveness}
After the first CAP, Test Phase 2 begins with the repaired vehicles. To capture
these revisions, the PM2 reliability growth model \cite{EH06} is often used.
This model explicitly captures testing phases, choices about which failure modes
to correct, and the potential of not completely eliminating a failure upon
repair. One of the downsides of PM2 is that many parameters of potential
interest, such as the Fix Effectiveness Factor (FEF), which measures how much
repairs improve failure rates, are typically fixed. A common value for FEF is
0.70. We follow the premise of this type of model, but allow a more flexible and
data-driven result that is less dependent on hard-coded assumptions.

One normal assumption used in reliability growth modeling is nondecreasing
failure rates; that is either the fixes were effective or had no effect, but did
not degrade the family of vehicles.  This should generally be the case, but
because we are dealing with complex systems and testing conditions are rarely
completely consistent, we will sometimes see decreases in failure rates after
adjustments are made.  Thus we will choose a prior distribution for the
correction parameter $\rho_{j}^{P2}$ that has positive support.  If
$\rho_{j}^{P2}$ is less than one, this represents an improvement in reliability.
And the same thinking is applied to $\rho_{j}^{P3}$ for the correction period
between phases two and three.  With this structure we can use the posterior
distributions of the correction parameters to make infrence on the effectiveness
of the repairs and the join posterior distributions to evaluate the overall
system reliability within each phase.

\subsubsection{Distributional Assumption}
The next modeling choice that must be made is selecting an appropriate
probability model This includes a selecting a sampling distribution for the data
and prior distributions for each parameter on which the sampling distribution
depends.  For lifetime data there are a number sampling distribution to choose
from.  Some common choices are exponential, Weibull, lognormal, gamma, inverse
Gaussian, and normal failure time models.  Because all of these distributions
have rate parameter (sometimes called scale = 1/rate) they can be applied to
the proposed modeling structure.  For this paper, we will focus on the
exponential and Weibull distributions.

First the exponential model,
\begin{equation*}
    f(y_{ijk}|\lambda_{ij})=\lambda_{ij} \exp(-\lambda_{ij}y_{ijk})
\end{equation*}

This is by far the most common parametric distribution used in reliability
modeling because of its desirable math mathematical properties and simple
interpretations.  One of these properties is that the hazard rate is constant.
For the JLTV, setting this would assume that a given component is equally likely
to fail in the first mile of testing as it is in the thousandth.  This may be
reasonable for some reliability settings, but in most setting we observe either
a decreasing hazard rate where failures are likely to occur early in testing
but less likely as the test goes on, or increasing hazard rate where we see the
system wear down as the test goes on.

Despite its common uses the assumption of a constant hazard rate over time is
rarely justifiable.  It has been well documented the issues that can arise when
this assumption is violated ~\cite{ref2}.  Thus we will also consider the same
hierarchical model structure while using the Weibull distribution.  Now we
assume that the failure mileage $t_{ijk}$ each follow a $Weibull(\lambda, \gamma)$
distribution with scale parameter $\lambda$ and shape parameter $\gamma$, with
probability density function,
\begin{equation*}
    f(t_{ijk}|\lambda_{ij},\gamma_{i})=\lambda_{ij}\gamma_{i}
    t_{ijk}^{\gamma_{i}-1}\exp(-\lambda_{ij} t_{ijk}^{\gamma_{i}}).
\end{equation*}
The Weibull distribution is a more flexible model with both a rate parameter
$\lambda_{ij}$ and a shape parameter $\gamma_{j}$.  The exponential is a special
case of the Weibull, when $\gamma = 1$.  You should notice here we are now
modeling the actual failure mileage $t_{ijk}$ instead of the miles between
failures $y_{ijk}$.  This means we are treating all intra-test repairs as "bad
as old".  We could have continued with the miles between failures and assumed
"good as new" repairs, but "bad as old" modeling will be more convenient for
developing assurance test plans in later sections.  This distinction is
important in the Weibull case because the failure rate is not constant in time.
In practice the decision on intra-test repair assumptions should be made based
on the application.  In many cases repairs should be treated as somewhere
between "good as new" and "bad as old".  This requires more complex modeling
that is beyond the scope of this paper.

It should  also be noted that we are currently indexing both $\rho$ and
$\gamma$, only by $j$ and not $i$, in other words, we are assuming a single
shape and between phase adjustment parameter for each failure mode across
vehicles.  All of these parameters could also be indexed by $i$ and modeled
hierarchical or with some combination of both.  This is something we will omit
for this paper but hope to explore in the future.

\subsubsection{JLTV Reliability}
So far we have discussed a number of different models and the assumptions that
accompany them.  Using McMC we can fit each of these models to the JLTV dataset.
Then we can use the posterior distributions of the systems parameters to
estimate reliability-related quantities of interest for each of the eight
vehicle types.  We are able to look at a number of different statistics related
to reliability at both  the component and system level.  In this study our main
quantity of interest is expected miles to  failure.  In the following figure we
have box plots showing the posterior distributions of  the expected miles to
failure for one of the vehicle types for each of the three phases, first using
the exponential model and second using the more general Weibull model.

% going to add a table that will have the posterior means and 90% credible
% interval limits
\begin{figure}
  \centering
      \includegraphics[width=8cm, height=6cm]{Boxplots}
  \caption{Comparing EMTF for Different Models}
\end{figure}

Figure 1 show that the posterior distributions from the Weibull model have a
good amount more spread than the exponential.  This is somewhat expected given
the Weibull model is estimating two model parameters to the exponentials one,
but what is concerning is that the center of the distributions is also quite
different between the two models.  All of the Weibull EMTF posterior means
are higher than the exponentials.  This is the first sign that the exponential
model may not be a good fit to the data.

\pagebreak

\subsubsection{Model Diagnostics}
As seen in the last section with the JLTV data, the decision of which model to
use can have a drastic impact on reliability assessment and future test
planning.  We will now look at a few  goodness-of-fit methods that can be used
to help in this decision.  The first model selection question we will explore is
the parametric form, exponential versus Weibull.  In statistical modeling we
often face the trade-off between fit and interpretability, and this situation is
no different. Because the Weibull is a more flexible model, it will always, in a
sense, fit the data better, but this comes at a price.  The exponential's
convenient form makes both computation and interpretation straightforward.  When
the Weibull's shape parameter is introduced this advantage is lost.  Thus, when
the overall fit is close to the same between the exponential and the Weibull we
will default to using the exponential.

The first check we used to decide between the exponential and Weibull models is
to fit the Weibull model and look at the posterior distributions of the
shape parameters and determine if one is a reasonable value.  Looking at the
plot below from the JLTV data, we see that for the 26 components it appears
that the value of one falls in the extreme tails of the distributions.  This is
our first clue that the exponential model will not be a good fit to this data.

% this plot will be remade without title and smaller xlim
\begin{figure}
  \centering
      \includegraphics[width=6cm, height=5cm]{betaPostPlot}
  \caption{Posterior Plots of Shape Parameter}
\end{figure}

This is not a surprising result, because the exponential model assumes constant
failure rate.  For most vehicle components we would expect the failure rate to
increase as the distance driven grows larger, and this corresponds with a shape
parameter between zero and one.  On the other hand, when the shape parameter is
larger than 1, the failure rate will decrease with time/distance.

Now a more formal diagnostic tool, the Deviance information criterion (DIC).
This is a popular method for comparing the goodness of fit of multiple models.
The DIC method gives a slight penalty for larger numbers of parameters.  The DIC
value is a unitless measure with lower values indicating a better fit to the
data.  In the table below we show the DIC results for the different distribution
and structure combinations, we considered with the JLTV dataset.

% need to recheck these numbers %
\begin{tabular}{|l|l|r|}
\multicolumn{3}{c}{\textbf{Goodness of Fit}} \\
\cline{1-3}
Distribution    & Structure & DIC \\
\hline
Exponential   & Common Rate Model                     & 21,264.19             \\
              & Hierarchical Rate Model               & 21,244.47             \\
Weibull       & Common Rate Model                     & 20,658.25             \\
              & Hierarchical Rate Model               & \textbf{17,908.06}    \\
\hline
\end{tabular}

Here we have shown that the hierarchical Weibull model fits the data much better
than the exponential, but in general information criteria comparisons do not
tell us if the model is truely a good representation of the underlying process.
The last method we will present is called posterior predictive checking
~\cite{ref3}. Here we are given important features of the data. In the JLTV case
we were interested in the total failure counts for each phase and the number of
time the miles between failures was less than 140.  We then used the final model
to simulate 5,000 new datasets. We then plot histograms for each of the 8
vehicles for all 3 phases.  In Figure 1 are examples of the histograms produced,
with the line showing where the value of the true dataset fell.  For this method
we don’t expect all of the true values to fall in the center of the
distribution.  Values in the tales are to be expected in a random processes like
this one.  We will only be concerned if we find true values falling in the
extreme tails or if we see a common bias to the high or low side of the
distributions.

% these plots will be redone and I plan to add a contrasting version of them
% using the exponential model
\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.4\linewidth]{Rplot1}
  \caption{Plot 1}
  \label{fig:sub1}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.4\linewidth]{Rplot2}
  \caption{Plot 2}
  \label{fig:sub2}
\end{subfigure}
\caption{Posterior Predictive Plots}
\label{fig:test}
\end{figure}

\subsection{Assurance Test Planning}
In this section, we will use the reliability models fit in section two section
to develop a plan for future testing.  Here the objective is to demonstrate that
at a desired level of confidence, the system will meet or exceed a specified
requirement. The methods to be employed will be similar to the assurance testing
as discussed by Hamada et al. ~\cite{ref4}. Bayesian \emph{reliability assurance
tests} are used to ensure that the reliability of an item meets or exceeds a
specified requirement with a desired probability. Meeker and Escobar
~\cite{ME04} established the difference between this and the more traditional
\emph{reliability demonstration test}.  The difference being that the assurance
test plan incorporates supplementary data and information, where the
demonstration test only uses test data.

In both test planning methods we are interested in controlling error rates while
minimizing the resources required for testing.  In the DoD acquisition process,
the error rates will be referenced in terms of risk. \emph{Consumer risk} which
considers the event of purchasing a product that does not meet reliability
requirements.  \emph{Producer risk} which considers the event of a product with
acceptable reliability failing a given test and not being considered for
purchase.  Introducing the notation for this section, in the JLTV case study, we
will be determining how many miles on test $T$, are required and how many system
failures to allow $c$, before the product is considered unacceptable.  We will
define $W(t)$ as a random variable that represents the total number of system
failures after $t$ miles.  We will let $\pi$ denote some quantity of interest
that is related to system reliability at a given time. It is common to base both
classical and Bayesian test plans on two specified levels of $\pi$: $\pi_0$, an
\emph{acceptable reliability level} (ARL), and $\pi_1$, a \emph{rejectable
reliability level} (RRL), where $\pi_1 \leq  \pi_0$. Although the precise
definition of ARL and RRL differ between the classical and Bayesian test
criteria, we use them in an equivalent way.

\subsubsection{Classical Approach}
In the frequentist (classical) version we will use two criteria in determining
the test plans. The \emph{producer's risk}, defined as the probability of
failing the test when $\pi = \pi_0$, and the \emph{consumer's risk}, defined as
the probability of passing the test when $\pi = \pi_1$.  Suppose that we specify
a maximum value, $\alpha$, of the producer's risk and a maximum value, $\beta$,
of the consumer's risk. For test planning, these criteria become
$$
\begin{aligned}
	\text{Producer's Risk} &= P \text{(Test Is Failed } \vert \pi_0 \text{)} \\ &=
	P \text{(} W(t) > c \vert \pi_0 \text{)} \leq \alpha
\end{aligned}
$$
and
$$
\begin{aligned}
	\text{Consumer's Risk} &= P \text{(Test Is Passed } \vert \pi_1 \text{)} \\ &=
	P \text{(} W(t) \leq c \vert \pi_1 \text{)} \leq \beta
\end{aligned}
$$
To choose a test plan for specified values of $(\alpha, \pi_0, \beta, \pi_1)$,
we assume a distributional form that defines the relationship between the number
of system failyres $W(t)$ and the reliability $\pi$.  Then we can simply find
the combinations of these probabilities by simultaneously solving these two
equations.  In the case where the component failure times are assumed to be
exponentially distributed with rate parameter $\lambda_i$ the failure counts for
each component come from a homogeneous Poisson process $w_i(t) \sim
\text{Poisson}(\lambda_i t)$, and the system failure counts $W(t) \sim
\text{Poisson}(\lambda_S t)$.  Numerous textbooks provide additional details of
this purely classical approach.

\subsubsection{Bayesian Approach}
Traditional methods for determining the testing procedure only rely on
distributional assumptions and/or asymptotic results. With a Bayesian approach we
incorporate the supplementary data from the previous testing phases with the
hopes of minimizing the resources needed for testing.

We now consider fully Bayesian posterior risks that convey a completely
different outlook from the corresponding classical risks. While the classical
provide assurance that satisfactory devices will pass the test and that
unsatisfactory devices will fail it, posterior risks provide precisely the
assurance that practitioners often desire: if the test is passed, then the
consumer desires a maximum probability $\beta$ that $\pi \leq \pi_1$. On the
other hand, if the test is failed, then the producer desires a maximum
probability $\alpha$ that $\pi \geq \pi_0$. Unlike the average risks, these
posterior risks are fully Bayesian in the sense that they are subjective
probability statements about $\pi$.

For a test that fails, the \emph{posterior producer's risk} is the probability
that $\pi \geq \pi_{0}$, or $P \text{(}\pi \ge \pi_0 \vert \text{Test Is
Failed)}$. Notice that this is simply the posterior probability that $\pi \ge
\pi_0$ given that we have observed more than $c$ failures. In the exponential
case, if we let $\pi$ be the system failure rate $\lambda_S$, then using Bayes'
Theorem, and assuming a maximum allowable posterior producer's risk $\alpha$, an
expression for the posterior producer's risk for the exponential test plan
$(T,c)$ is
$$
\begin{aligned}
    P(\lambda_S \geq \lambda_0 \; \vert \; \text{Test Is Failed)} &= P(\lambda_S
    \geq \lambda_0 \; \vert \; W > c) \\ &= \int_{\lambda_0}^{\infty}
    p(\lambda_S \; \vert W > c) d\lambda_S \\ &= \int_{\lambda_0}^{\infty}
    \frac{f(W > c \vert \lambda_S)}{\int_{0}^{\infty} f(W > c \vert \lambda_S)
    d\lambda_S} d\lambda_S \\ &= \frac{\int_{\lambda_0}^{\infty} 1 - [ \sum_{W=0}^c
    \frac{(\lambda_S T)^W exp(-\lambda_S T)}{W!}]p(\lambda_S)d\lambda_S}
    {\int_{0}^{\infty} 1 - [ \sum_{W=0}^c \frac{(\lambda_S T)^W exp(-\lambda_S
    T)}{W!}]p(\lambda_S)d\lambda_S} \leq \alpha
\end{aligned}
$$
\\
\\
For simplicity we if fix c to be zero and then perform Monte Carlo integration
using N posterior draws $ \lambda_S^{(j)} $
$$
\begin{aligned}
	 P(\lambda_S \geq \lambda_0 \; \vert \; W = 0) &=
	 \frac{\int_{\lambda_0}^{\infty} 1 - exp(-\lambda_S T)p(\lambda_S)d\lambda_S}
	 {\int_{0}^{\infty} 1 - exp(-\lambda_S T)p(\lambda_S)d\lambda_S} \\ &\approx
	 \frac{\sum_{j = 1}^{N} 1 - exp(-\lambda_S^{(j)} T)I(\lambda_S^{(j)} \geq
	 \lambda_0)} {\sum_{j = 1}^{N} 1 - exp(-\lambda_S^{(j)} T)} \end{aligned}
$$

Similarly, given that the test is passed, the \emph{posterior consumer's risk}
is the probability that $\pi \leq \pi_1$, or $P \text{(}\pi \leq \pi_0 \vert
\text{Test Is Passed)}$. Notice that this is simply the posterior probability
that $\pi \leq \pi_1$ given that we have observed no more than $c$ failures. In
the exponential case, if we let $\pi$ be the system failure rate $\lambda_S$,
then using Bayes' Theorem, and assuming a maximum allowable posterior consumer's
risk $\beta$, an expression for the posterior producer's risk for the
exponential test plan $(T,c)$ is

$$
\begin{aligned}
    P(\lambda_S \leq \lambda_1 \; \vert \; \text{Test Is Passed)} &= P(\lambda_S
    \leq \lambda_1 \; \vert \; W \leq c) \\ &= \int_{0}^{\lambda_1} p(\lambda_S \;
    \vert W > c) d\lambda_S \\ &= \int_{0}^{\lambda_1} \frac{f(W > c \vert
    \lambda_S)}{\int_{0}^{\infty} f(W > c \vert \lambda_S) d\lambda_S} d\lambda_S \\ &=
    \frac{\int_{0}^{\lambda_1} [ \sum_{W=0}^c \frac{(\lambda_S T)^W
    exp(-\lambda_S T)}{W!}]p(\lambda_S)d\lambda_S} {\int_{0}^{\infty} [
    \sum_{W=0}^c \frac{(\lambda_S T)^W exp(-\lambda_S
    T)}{W!}]p(\lambda_S)d\lambda_S} \leq \beta
\end{aligned}
$$
\\
\\
For simplicity we if fix c to be zero and then perform Monte Carlo integration
using N posterior draws $ \lambda_S^{(j)} $ $$
\begin{aligned}
	 P(\lambda_S \leq \lambda_1 \; \vert \; W = 0) &= \frac{\int_{0}^{\lambda_1}
	 exp(-\lambda_S T)p(\lambda_S)d\lambda_S} {\int_{0}^{\infty} exp(-\lambda_S
	 T)p(\lambda_S)d\lambda_S} \\ &\approx \frac{\sum_{j = 1}^{N}
	 exp(-\lambda_S^{(j)} T)I(\lambda_S^{(j)} \leq \lambda_1)} {\sum_{j = 1}^{N}
	 exp(-\lambda_S^{(j)} T)}
\end{aligned}
$$

\subsubsection{Weibull Case}
In the exponential case, we defined our quantity of interest that is related to
system reliability at a given time $\pi$ to be equal to the system failure rate
$\lambda_S$.  This distributional assumption gave us a convenient distributional
form for our system failure counts with $W(t) \sim \text{Poisson}(\lambda_S t)$
When we assume the miles to failure for each system component follows a Weibull
distribution the system failure counts no longer follow a homogeneous Poisson
process.  When assuming "bad as old" intra-test repairs we do find that the
failure count for a given failure mode does follow a nonhomogeneous Poisson
process, $w(t)_i \sim \text{Poisson}( (\lambda_i t)^{\gamma_i} )$.  In this case
we no longer have a closed form for the system failure rate $\lambda_S = \sum_i
(\lambda_i t)^{\gamma_i}$.  Because of this, using our previous approach leads to
an unpleasant series of integals with parameters in the bounds.  Thankfully we
can once again lean on Monte Carlo integration and the expression for the
posterior producer's risk for the Weibull test plan $(T,c)$ simplifies to
$$
\begin{aligned}
	 P(\lambda_S \geq \lambda_0 \; \vert \; W = 0) &= \frac{\int_{0}^{\infty}
	 \int_{\lambda_0}^{\infty} 1 - exp(- \gamma \lambda_S T^{\gamma})p(\lambda_S)
	 p(\gamma) d\lambda_S d\gamma} {\int_{0}^{\infty} \int_{0}^{\infty} 1 - exp(-
	 \gamma \lambda_S T^{\gamma})p(\lambda_S) p(\gamma) d\lambda_S d\gamma} \\
	 &\approx \frac{\sum_{j = 1}^{N} 1 - exp(-\gamma^{(j)} \lambda_S^{(j)}
	 T^{\gamma^{(j)}}) I(\lambda_S^{(j)} \geq \lambda_0)} {\sum_{j = 1}^{N} 1 -
	 exp(- \gamma^{(j)} \lambda_S^{(j)} T^{\gamma^{(j)}})}
\end{aligned}
$$
\\
Once again set c to be zero and used N posterior draws for both
$\lambda_S^{(j)}$ and $\gamma^{(j)}$

It should be noted that in this test planning process you would use the
posterior distribution of $\lambda_S$ from the final developmental testing
phase.  It is also common to incorporate a degradation factor to account for an
expected reduction in reliability from the developmental phase to the
operational phase.  In the Bayesian framework, this entails adding another
multiplicative parameter to the rate parameter with an appropriate prior.   A
10-30 percent reduction from the developmental phase to the operational phase is
common in many DoD applications.

\subsubsection{JLTV Test Plan Results}
The first step in applying this method to our JLTV dataset is defining $\pi_0$,
$\pi_1$, $\alpha$ and $\beta$.  That is the ARL, RRL, acceptable producer risk
and acceptable consumer risk respectively.  With input from subject matter
experts  we will define the producer's target or ARL as the expected number of
failures in the first 140 miles to be less  than 2, and the consumer's minimum
requirements or RRL as the expected number of failures in  the first 80 miles to
greater than 2.  We will use traditional values of 0.1 and 0.05 for $\alpha$ and
$\beta$ respectively, but in practice these values should always be set based on
the true risk tolerances of the consumer and producers.  This gives us the
following probability statements to satisfy when designing the test:
\\
Consumer Risk : $ P(80^\beta (\lambda_S) \geq 2 \; \vert \text{Test Is Passed}) \leq 0.1 $ \\
Producer Risk : $ P(140^\beta (\lambda_S) \leq 2 \; \vert \; \text{Test Is Failed}) \leq 0.05 $

From here we are able to find combinations of failures allowed and test miles
required that satisfy both risk requirements.  The table below show the
resulting combinations for the JLTV data.

\begin{tabular}{|l|l|r|}
\multicolumn{2}{c}{\textbf{JLTV Test Plan}} \\
\cline{1-2}
Failures Allowed    & Test Miles Required \\
\hline
1   & 5,300      \\
2   & 6,500      \\
3   & 7,800      \\
4   & 11,100     \\
5   & 13,100     \\
\hline
\end{tabular}

% This part should be expanded quite a bit and will extend into the discussion
% section.
It should be noted that if $\pi_0$, $\pi_1$, $\alpha$ and $\beta$ are set to
unrealisitc values there can be situations where no combinations of failures
allowed and test miles required will satisfy these probability statements.  In
this paper we have focused on test plans for single units, but this methodology
can be extended to find test plans that allow for a changing number of testing
units.  In this case you would find combinations of failures allowed and test
miles required and number of vehicles tested that satisfy the risk requirements.
More details and examples of this can be found in the assurance testing section
of Hamada et al. ~\cite{ref4}

\subsection{Discussion}

%%%%% ref %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{thebibliography}{99}
\bibitem{ref1} Director, Defense Test and Evaluation. Test and evaluation of
system reliability availability maintainability - A primer; Third Edition
(1982).
\bibitem{ref2} National Research Council. Statistics, Testing, and Defense
Acquisition: Background Papers. \textit{National Academies Press} (1999).
\bibitem{ref3} Gelman, Andrew, Xiao-Li Meng, and Hal Stern. "Posterior
predictive assessment of model fitness via realized discrepancies."
\textit{Statistica sinica} 733-760 (1996).
\bibitem{ref4} M.\ S.\ Hamada, A.\ G.\ Wilson, C.\ S.\ Reese and H.\ F.\ Martz.
Bayesian Reliability. Springer (2008).
\bibitem{ME04} W.\ Q.\ Meeker and L.\ A.\
Escobar. Reliability: the other dimension of quailty. \textit{Quality Technology
and Quantitative Management} 1, 1-25 (2004). \bibitem{HWWGM13} M.\ S.\ Hamada,
A.\ G.\ Wilson, B.\ P.\ Weaver, R.\ W.\ Griffiths and H.\ F.\ Martz. Bayesian
binomial assurance tests for system reliability using component data.
\textit{Journal of Quality Technology} 46, 24-32 (2014). \bibitem{JGHR03} V.\
E.\ Johnson, T.\ L.\ Graves, M.\ S.\ Hamada, and C.\ S.\ Reese. A hierarchical
model for estimating the reliability of complex systems. In Bayesian Statistics,
Vol. 7,\ J.\ M.\ Bernardo, M.\ J.\ Bayerri, J.\ O.\ Berger, A.\ P.\ Dawid, D.\
Heckerman, A.\ F.\ M.\ Smith, and M.\ West eds., Oxford, UK: Oxford University
Press (2003). \bibitem{HMRGJW04} M.\ S.\ Hamada, H.\ F.\ Martz, C.\ S.\ Reese,
T.\ L.\ Graves, V.\ E.\ Johnson, and A.\ G.\ Wilson. A fully Bayesian approach
for combining multilevel failure information in fault tree quantification and
optimal follow-on resource allocation. \textit{Reliability Engineering and
System Safety} 86, 397-405 (2004). \bibitem{EH06} P.\ M.\ Ellner and J.\ B.\
Hall, An approach to reliability growth planning based on failure mode discovery
and correction using AMSAA projection methodology. \textit{IEEE Proceedings of
the Annual Reliability and Maintainability Symposium} (2006).
\end{thebibliography}

\section{Appendix A1}
\textbf{Exponential Model:}
\begin{equation*}
L(\lambda_1, ..., \lambda_8, \rho^{P2}, \rho^{P3}, \vert Y) = \prod_{i = 1}^8
  \Bigg[\prod_{j = 1}^{n_j}(\lambda_i p^*e^{-\lambda_i p^* y_{ij}})\prod_{j =
  1}^R(e^{-\lambda_i p^*
  y_{ij}})\Big(\frac{\beta^\alpha}{\Gamma(\alpha)}\lambda_i^{\alpha -
  1}e^{-\beta\lambda_i}\Big)\Bigg]
\end{equation*}
\[p^* =   \left\{
\begin{array}{ll}
      1 & y_{ij}\in \text{Phase 1} \\
      \rho_{j}^{P2} & y_{ij} \in \text{Phase 2} \\
      \rho_{j}^{P2} \rho_{j}^{P3} & y_{ij} \in \text{Phase }
\end{array}
\right. \]
$R$ is the number of right-censored observations, and n is the number of
non-censored observations.

\begin{equation*}
\pi(\alpha, \beta, \rho^{P2}, \rho^{P3}) \propto (\alpha^{a_1-1}e^{-\alpha
  a_2})(\beta^{b_1 - 1}e^{-\beta b_2})(\rho^{P2*c_1 - 1}e^{-\rho^{P2}
  c_2})(\rho^{P3*d_1 - 1}e^{-\rho^{P3} d_2})
\end{equation*}

\begin{align*}
\lambda_i|\cdot &\sim \text{Gamma}(n_i + \alpha, \sum_{j=1}^{n_i}y_{ij}p^* + \beta) \\
\beta|\cdot &\sim \text{Gamma}(8\alpha + b_1, \sum{\lambda_i} + b_2) \\
\rho^{P2}|\cdot &\sim \text{Gamma}(N_{1,2} + c_1, \sum{y_{ij}\lambda_i} + c_2) \\
\rho^{P3}|\cdot &\sim \text{Gamma}(N_{2} + d_1, \sum{y_{ij}\lambda_i} + d_2) \\
\alpha|\cdot &\sim \text{Unknown form.  Used Metropolis-Hastings.}
\end{align*}

$N_1$ and $N_2$ are the total non-censored observations for phases two and three
and phase three, respectively.

\textbf{Weibull Model (Bad as Old):}
\begin{equation*}
  L(\lambda_1, ..., \lambda_8, \gamma, \rho^{P2}, \rho^{P3}, \vert T) = \prod_{i =
    1}^8 \Bigg[\prod_{j = 1}^{n_j}(\lambda_i p^*t_{ij}^{\gamma_i-1} e^{-\lambda_i
    p^* t_{ij}^{\gamma_i}})\prod_{j = 1}^R(e^{-\lambda_i p^*
    t_{ij}^{\gamma_i}})\Big(\frac{\beta^\alpha}{\Gamma(\alpha)}\lambda_i^{\alpha -
    1}e^{-\beta\lambda_i}\Big)\Bigg]
\end{equation*}
\[p^* =   \left\{
\begin{array}{ll}
      1 & t_{ij}\in \text{Phase 1} \\
      \rho^{P2^\gamma} & t_{ij} \in \text{Phase 2} \\
      \rho^{P2^\gamma} \rho^{P3^\gamma} & t_{ij} \in \text{Phase }
\end{array}
\right. \]
$R$ is the number of right-censored observations, and n is the number of
non-censored observations.

\begin{equation*}
\pi(\alpha, \beta, \rho^{P2}, \rho^{P3}) \propto (\alpha^{a_1-1}e^{-\alpha
  a_2})(\beta^{b_1 - 1}e^{-\beta b_2})(\rho^{P2*c_1 - 1}e^{-\rho^{P2}
  c_2})(\rho^{P3 * d_1 - 1}e^{-\rho^{P3} d_2})(\gamma^{g_1 - 1}e^{-\gamma g_2})
\end{equation*}

\begin{align*}
\lambda_i|\cdot &\sim \text{Gamma}(n_i + \alpha, \sum_{j=1}^{n_i}t_{ij}^{\gamma_i} p^* + \beta) \\
\beta|\cdot &\sim \text{Gamma}(8\alpha + b_1, \sum{\lambda_i} + b_2) \\
\rho^{P2}|\cdot &\sim \text{Gamma}(N_{1,2} + c_1, \sum{t_{ij}^{\gamma_i} \lambda_i} + c_2) \\
\rho^{P3}|\cdot &\sim \text{Gamma}(N_{2} + d_1, \sum{t_{ij}^{\gamma_i} \lambda_i} + d_2) \\
\alpha|\cdot &\sim \text{Unknown form.  Used Metropolis-Hastings.}\\
\gamma|\cdot &\sim \text{Unknown form.  Used Metropolis-Hastings.}
\end{align*}

$N_1$ and $N_2$ are the total non-censored observations for phases two and three
and phase three, respectively.

\end{document}